Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-05-31 fib2.tex
edits. add a ref.

\usepackage{tikz}
\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,positioning,fit}
\usepackage{epstopdf} % to include .eps graphics files with pdfLaTeX
%\usepackage{bm}  % Define \bm{} to use bold math fonts

@@

@@

We use a phenomenological model of Fibonacci anyon dynamics which neglects any 
microscopic details of the system. 
This is consistent with the principles of topologically 
ordered systems and anyonic physics, where the key universal features describing the 
anyon model correspond to large length-scale physics, while the microscopic physics plays 
a less important (and non-universal) role. Effectively, we simulate a $(2+1)$-dimensional topological quantum field theory on a discrete space-time.
As such, we model the system simply as a 
periodic $L\times L$ square lattice $\Lambda$ of sites upon which sets of anyons can 
reside. Allowed anyon dynamics (pair-creation, braiding, and fusion) are considered to act 
locally on edges of $\Lambda$. For more details of an analogous model, see~\cite{Brell2013}.

\dude{I really need to see something about observables vs. states. Even if
we dont mention tqft.}

@@

Our error-correction algorithm is based on a hard-decision renormalization group (RG) 
decoder~\cite{Bravyi2011}. 
The decoder proceeds by measuring the occupations of 
anyons at each site of the lattice, forming clusters of nearby anyons and fusing the anyons 
within each cluster, before iterating at linearly increasing length scales.
It terminates when the lattice 
is free of anyons (i.e.~we have returned to the codespace) or the clustering length scale exceeds the size of the lattice.
%the decoder performs a homologically nontrivial operation, thereby altering the codespace.
%\cggb{our simulation may terminate when the decoder performs a nontrivial operation, but here we're just describing the decoder, which can't know that.}
%\cggb{A few more details? What is our length schedule?} 
%\cggb{How does our decoder relate to other decoding schemes used in~\cite{Wootton2013,Brell2013,Hutter2014}?} 
Note that, as was found in Ref.~\cite{Brell2013}, we expect that our qualitative results may be reproduced by most alternative families of decoders, although it may not be clear in all cases how to extend these decoders to the non-abelian setting.
The advantage of using a hard-decision RG decoder is 
its simplicity and flexibility, and the fact that its clustering scheme is compatible with the structure in the noise processes that allows us to classically simulate them.

@@

Below the (bond) percolation threshold for (say) a 2D square lattice, we expect random sets of 
bonds to decompose into separate clusters~\footnote{We use the word cluster to refer both to groupings of nearby anyons as formed by the decoding algorithm and to connected components of a graph under percolation, since both of these usages are standard in the literature. The intended meaning will remain clear from context.} of average size $O(\log(n))$ and variance 
$O(1)$~\cite{Bazant2000}.
Each noise process in our model is associated with a (randomly distributed) edge, and so 
disconnected clusters correspond to sets of anyons that could not have interacted at any 
point in their history. Although normally anyons can interact by braiding without ever coming into contact with one another, in this case charge conservation implies that each distinct cluster must have trivial total charge, and so all such interactions are trivial.
This allows us to simulate the braiding processes within each cluster 
separately. In other words: the quantum state in the fusion space of all anyons factorizes into 
a tensor product over bond clusters, and we can compute each factor separately. Since each 
cluster has size only $O(\log(n))$, we can typically simulate these dynamics efficiently 
because the resulting fusion space has dimension $O(\mathrm{poly}(n))$. There are rare 
random processes that violate this reasoning, but processes such as these are suppressed 
exponentially in the lattice size $L$\cite{Grimmett1989}.

@@

Of course, if we were to consider noise so strong that the clusters percolate over the entire lattice, we would no longer expect to be able to efficiently simulate the noise and recovery processes in the system. 
However, recall that logical errors in our system correspond to processes that act over a non-trivial loop of the torus. 
Our simulation remains efficient until the point that we start to create nontrivial loops around the torus, and our simulation begins to declare failure whenever non-trivial logical operations occur, meaning that we can simulate large lattice sizes in the successful error-correction (sub-percolated) regime, including up to $L=128$.

@@

Our concrete simulation proceeds by associating a \emph{tile} with each site of the 
$L\times L$ periodic square lattice.
In order to keep track of the observables of the system 
we maintain a set of disjoint curves, representing the history of anyons on the lattice.
These curves can also be thought of as a top-down view of the fusion tree:
the anyons comprising the fusion tree are here displayed as points along each curve. 
In general, the curves wind around the two 
dimensional tiles in a haphazard way determined by the progress of the simulation.
This allows a dynamically generated basis for the fusion space that easily allows for tensor 
decomposition into factors corresponding to disjoint clusters.
For comparison, previous work 
used a single fixed basis for the fusion space (determined by row-major ordering of 
sites)~\cite{Brell2013}.

\begin{figure}[th!]
\begin{center}
	\includegraphics[width=\columnwidth]{pic-syndrome.pdf}
\caption{Four equivalent representations of a state,
along with two syndrome observables marked with solid and dashed lines:
(a) two pairs are joined in (b).
Deform the curve straight to get (c),
then effect some
braid operations on the state so that (d) the two
observables are now represented on the curve. \cggb{I'm not totally clear on what I'm looking at here. Is this the diagram where we want to explain the whole simulation algorithm? Is the idea that we start with a), which is given by the set of noise processes that have happened, then we connect stuff up arbitrarily into b) to make some fusion tree. Then c) is what's happening after linearizing the picture? And d) is c) after some R and F moves so you go into the standard basis for the fusion tree? Are the red lines between c) and d) just denoting leaves with the same identity or do they have some more precise meaning wrt the braid moves that relate c) and d)?}}
\label{f:syndrome}
\end{center}
\end{figure}

\begin{figure}[th!]
\begin{center}
	\includegraphics[width=\columnwidth]{pic-decode.pdf}
\caption{The decoder works by maintaining a set of disjoint clusters (rooted trees). \cggb{Is this diagram where we want to explain the decoding algorithm? Is part a) round one of clustering and part b) round 2 of clustering? Do the arrow directions just denote the direction stuff goes when we try to fuse it all together or does it have some deeper meaning?}}
\label{f:decode}
\end{center}
\end{figure}

>>>>

%\usepackage{tikz}
%\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,positioning,fit}

\usepackage{epstopdf} % to include .eps graphics files with pdfLaTeX
%\usepackage{bm}  % Define \bm{} to use bold math fonts

@@

\dude{these equations are too big... i was hoping for something tight, like
in http://arxiv.org/abs/1504.02868
any ideas? i might just generate a bunch of little pdf's...
}

@@

We use a phenomenological model of Fibonacci anyon dynamics which neglects any 
microscopic details of the system. 
This is consistent with the principles of topologically 
ordered systems and anyonic physics, where the key universal features describing the 
anyon model correspond to large length-scale physics, while the microscopic physics plays 
a less important (and non-universal) role.
We simulate a $(2+1)$-dimensional topological quantum field theory.
%Effectively, we simulate a $(2+1)$-dimensional topological quantum field theory on a discrete space-time.
%As such, we model the system simply as a 
%periodic $L\times L$ square lattice $\Lambda$ of sites upon which sets of anyons can 
%reside. 
%Allowed anyon dynamics (pair-creation, braiding, and fusion) are considered to act 
%locally on edges of $\Lambda$. 
%For more details of an analogous model, see~\cite{Brell2013}.
\dude{something about observables vs. states.}

@@

Our error-correction algorithm is
based on a heirarchical clustering algorithm~\cite{Hastie2009},
and has some 
similarities to the hard-decision renormalization group (RG) decoder~\cite{Bravyi2011}. 
The decoder proceeds by measuring the occupations of 
anyons at each site of the lattice, forming clusters of nearby anyons and 
%fusing the anyons within each cluster. 
measureing the total charge within each cluster.
Clusters with trivial charge are discarded, and
the rest are joined (agglomeratively~\cite{Hastie2009})
at linearly increasing length scales.
See FIG. \ref{f:decode}.
Note that unlike the abelian case, the decoder
cannot know the charge of each cluster based only
on the syndrome information (as in \cite{Bravyi2011}),
and so will query (measure) the system to find this.
The simulation proceeds in this way as a dialogue between decoder
and system, terminating when the lattice 
is free of anyons (i.e.~we have returned to the codespace) 
or the decoder performs a homologically nontrivial operation, thereby altering the codespace.
%or the clustering length scale exceeds the size of the lattice.
%\cggb{our simulation may terminate when the decoder performs a nontrivial operation, but here we're just describing the decoder, which can't know that.}
%\dude{in our sim the decoder stops when there is one cluster. Ie., it never gives up.}
%\cggb{A few more details? What is our length schedule?} 
%\cggb{How does our decoder relate to other decoding schemes used in~\cite{Wootton2013,Brell2013,Hutter2014}?} 
Note that, as was found in Ref.~\cite{Brell2013}, we expect that our qualitative results may be reproduced by most alternative families of decoders, although it may not be clear in all cases how to extend these decoders to the non-abelian setting.
The advantage of using a hard-decision decoder is 
its simplicity and flexibility, and the fact that its clustering scheme is compatible with the structure in the noise processes that allows us to classically simulate them.

\begin{figure}[th!]
\begin{center}
	\includegraphics[width=1.0\columnwidth]{pic-decode.eps}
\caption{The decoder works by maintaining a set of disjoint clusters as rooted trees.
(a) Trees are formed from neighboring sites that contain a charge.
(b) For each successive round, we grow all trees, 
and overlapping trees are joined to form
a single tree. Arrows denote movement of charge to the root of the tree.
The root of each tree is chosen arbitrarily.
}
\label{f:decode}
\end{center}
\end{figure}

@@

Below the (bond) percolation threshold for (say) a 2D square lattice, we expect random sets of 
bonds to decompose into separate clusters~\footnote{
We use the word cluster to refer both to groupings 
of nearby anyons as formed by the decoding algorithm and 
to connected components of a graph under percolation\dude{confusing}, since both 
of these usages are standard in the literature. The intended 
meaning will remain clear from context.} of average size $O(\log(n))$ and variance 
$O(1)$~\cite{Bazant2000}.
Each noise process in our model is associated with a (randomly distributed) edge, and so 
disconnected clusters correspond to sets of anyons that could not have interacted at any 
point in their history. 
Although normally anyons can interact by braiding without ever coming 
into contact with one another \dude{um um um}, in this case charge conservation 
implies that each distinct cluster must have trivial total charge, and so all such interactions are trivial.
This allows us to simulate the braiding processes within each cluster 
separately. In other words: the quantum state in the fusion space of all anyons factorizes into 
a tensor product over bond clusters, and we can compute each factor separately. Since each 
cluster has size only $O(\log(n))$, we can typically simulate these dynamics efficiently 
because the resulting fusion space has dimension $O(\mathrm{poly}(n))$. There are rare 
random processes that violate this reasoning, but processes such as these are suppressed 
exponentially in the lattice size $L$\cite{Grimmett1989}.

@@

Of course, if we were to consider noise so strong that 
the clusters percolate over the entire lattice, we would no 
longer expect to be able to efficiently simulate the noise and recovery processes in the system. 
However, recall that logical errors in our system correspond to processes that act over a non-trivial loop of the torus. 
Our simulation remains efficient \dude{efficiency has nothing to do with it}
until the point that we start to create nontrivial loops around the torus, 
and our simulation begins to declare failure whenever non-trivial logical 
operations occur, meaning that we can simulate large lattice sizes 
in the successful error-correction (sub-percolated) regime, including up to $L=128$.

@@

Our concrete simulation proceeds by associating a \emph{tile} with each site of the 
$L\times L$ periodic square lattice.
In order to keep track of the observables of the system 
we maintain a set of disjoint curves, representing the history of anyons on the lattice.
These curves can also be thought of as a top-down view of the fusion tree:
the anyons comprising the fusion tree are here displayed as points along each curve. 
In general, the curves wind around the two 
dimensional tiles in a haphazard way determined by the progress of the simulation.
This allows a dynamically generated basis for the fusion space that easily allows for tensor 
decomposition into factors corresponding 
to disjoint clusters \dude{confusing use of this word}.
For comparison, previous work 
used a single fixed basis for the fusion space (determined by row-major ordering of 
sites)~\cite{Brell2013}.

\begin{figure}[th!]
\begin{center}
	\includegraphics[width=1.0\columnwidth]{pic-syndrome.eps}
\caption{Four equivalent representations of a state,
along with two syndrome observables marked with solid and dashed lines:
(a) two pairs are joined in (b).
Deform the curve straight to get (c),
then effect some
braid operations on the state so that (d) the two
observables are now represented on the curve. 
%\cggb{I'm not totally clear on what I'm looking at here. Is this the diagram where we want to explain the whole simulation algorithm? Is the idea that we start with a), which is given by the set of noise processes that have happened, then we connect stuff up arbitrarily into b) to make some fusion tree. Then c) is what's happening after linearizing the picture? And d) is c) after some R and F moves so you go into the standard basis for the fusion tree? Are the red lines between c) and d) just denoting leaves with the same identity or do they have some more precise meaning wrt the braid moves that relate c) and d)?}
}
\label{f:syndrome}
\end{center}
\end{figure}

In FIG. \ref{f:syndrome} we show two adjacent sites
involved in a syndrome measurement.
%This will be the total charge on each site, ie. either $\tau$ or $\mathbb{I}.$
The error process has resulted in two pair
creation events along the edge joining these sites.
\dude{i'd like to write out the equations for each of
these 4 steps, but we'd need to keep it tight somehow.}
First, (a) we show the state as a tensor product of two pairs $\tau\tau\otimes\tau\tau$.
Then, (b) these two pairs are embedded into a single larger fusion
tree $\tau\tau\tau\tau$.
This is continuosly deformable into the picture in (c).
It is now clear the two braid operations required so
that the observables are represented on the fusion tree.
The general algorithm and datastructures used
for this proceedure will be discussed elsewhere.

\dude{Arrgh, what do we say about how to simulate moving anyons around
the lattice ?}
